Modelling the impacts of replacing car trips with combined public transport and cycling: Reproducible methods, results and actionable evidence from the XXX project

### 2023-11-28 paper/PaperCEUS/PaperCEUS.Rmd
title and plot aml trips

---
title: "XXX: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_XXX_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 10% of trips could be made by bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the PT leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

@@

library(dplyr)
```

@@

According to the latest mobility survey conducted in 2018 [@IMOB], the LMA registered a total of 5.3 million daily trips, with only 0.5% by bicycle.
Car modal share was 58.4%, while PT accounted for 15.5%.
The number of intra-municipal trips --- with origin and destination in the same municipality --- amounts to 3.5 million trips.
This exceeds the number of inter-municipal trips (1.8 million trips), involving travel between different municipalities.
Cars and public transport are the most used modes for intercity trips, with cars being the predominant choice for all journeys.

<!-- Plot of  AML trips by mode and inter/intra %? -->

@@

This paper estimates the potential for combining cycling and PT to substitute car trips in the LMA.
After presenting the methods used, it assesses its socio-environmental impacts using open data and open-source tools.

>>>>

---
title: "Modelling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from the XXX project"
# subtitle: "Reproducible methods, results and actionable evidence from the XXX project"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_XXX_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 10% of trips could be made by bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the PT leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

@@

library(dplyr)
library(ggplot2)
```

@@

According to the latest mobility survey conducted in 2018 [@IMOB], the LMA registered a total of 5.3 million daily trips, with only 0.5% by bicycle.
Car modal share was 58.4%, while PT accounted for 15.5% (see Figure \ref{fig:mododist}).
The number of intra-municipal trips --- with origin and destination in the same municipality --- amounts to 3.5 million trips.
This exceeds the number of inter-municipal trips (1.8 million trips), involving travel between different municipalities.
Cars and public transport are the most used modes for intercity trips, with cars being the predominant choice for all journeys.

<!-- Plot of  AML trips by mode and inter/intra %? -->
```{r dadosaml, message=FALSE, warning=FALSE}
TRIPSmode_municipio = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/TRIPSmode_municipal.Rds"))

TRIPSmode_aml = TRIPSmode_municipio %>%
  group_by(modo) %>%
  summarise(viagens = round(sum(viagens))) %>% 
  mutate(percentagem = round(100*viagens/sum(viagens),1))

TRIPSmode_aml_intr = TRIPSmode_municipio
TRIPSmode_aml_intr$Inter[TRIPSmode_aml_intr$Origem == TRIPSmode_aml_intr$Destino] = "Intra-municipal"
TRIPSmode_aml_intr$Inter[is.na(TRIPSmode_aml_intr$Inter)] = "Inter-municipal"
TRIPSmode_aml_intr = TRIPSmode_aml_intr %>% 
  group_by(modo, Inter) %>%
  summarise(viagens = round(sum(viagens)))

viagensinter = sum(TRIPSmode_aml_intr$viagens[TRIPSmode_aml_intr$Inter == "Inter-municipal"])
viagensintra = sum(TRIPSmode_aml_intr$viagens[TRIPSmode_aml_intr$Inter == "Intra-municipal"])

TRIPSmode_aml_intr$percentagem[TRIPSmode_aml_intr$Inter == "Inter-municipal"] =  round(100*TRIPSmode_aml_intr$viagens[TRIPSmode_aml_intr$Inter == "Inter-municipal"]/viagensinter, 1)

TRIPSmode_aml_intr$percentagem[TRIPSmode_aml_intr$Inter == "Intra-municipal"] =  round(100*TRIPSmode_aml_intr$viagens[TRIPSmode_aml_intr$Inter == "Intra-municipal"]/viagensintra, 1)

TRIPSmode_aml_intr = bind_rows(TRIPSmode_aml %>% mutate(Inter = "All"),  TRIPSmode_aml_intr)

```
```{r mododist, fig.ncol=2, out.width="50%", fig.cap="\\label{mododist}Trips in the LMA by inter/intra municipal and mode, according to the travel survey.", fig.subcap=c("In percentage","In total")}
#gráfico com bistribuição de viagens por modo na AML segundo IMob 2018

TRIPSmode_aml_intr$modo = factor(TRIPSmode_aml_intr$modo, 
                                 levels = c("Car", "Transit", "Motorcycle", 
                                            "Other", "Walk", "Bike"))

col_modes = c("#fe5f55", "#ffd166", "#8E794A", "grey", "#457b9d", "#90be6d")

ggplot(TRIPSmode_aml_intr, aes(fill = modo, y = percentagem/100, x = Inter)) +
  geom_bar(position = position_fill(reverse = FALSE), stat = "identity") +
  scale_fill_manual(values = col_modes, name = "Mode") +
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(label = ifelse(percentagem > 6, paste0(percentagem,"%"), "")),
    position = position_stack(vjust = 0.5, reverse = FALSE))+
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "none") +
  coord_flip()

ggplot(TRIPSmode_aml_intr, aes(fill = modo, y = viagens, x = Inter)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = col_modes, name = "Mode") +
  scale_y_continuous("Trips / day", labels = scales::comma_format(big.mark = ' ')) +
  geom_text(aes(label = ifelse(viagens > 500000, paste0(round(viagens/1000000,1)," M"), "")),
  position = position_stack(vjust = 0.5, reverse = FALSE))+
  theme_minimal() +
  theme(
    axis.title.y = element_blank(),
        legend.position = "right") +
  coord_flip()

@@

This paper estimates the potential for combining cycling and PT to substitute car trips in the LMA, while achieving the national cycling targets.
After presenting the methods used, it assesses its socio-environmental impacts using open data and open-source tools.

### 2023-11-28 paper/PaperCEUS/PaperCEUS.tex
title and plot aml trips

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={XXX: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

  \title{XXX: Estimating the socio-environmental impacts of car
substitution by bicycle and public transit using open tools}
    \author[CERIS]{Rosa Félix\corref{cor1}%
  \corref{cor1}%
  }
   \ead{rosamfelix@tecnico.ulisboa.pt} 
    \author[CERIS]{Filipe Moura%
  %
  }
   \ead{fmoura@tecnico.ulisboa.pt} 
    \author[ITS]{Robin Lovelace%
  %
  }
   \ead{r.lovelace@uleeds.uk} 
      \affiliation[CERIS]{CERIS, Instituto Superior Técnico -
Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal}
    \affiliation[ITS]{Institute for Transport Studies, University of
Leeds. 34-40 University Rd, Leeds LS2 9JT, UK}
    \cortext[cor1]{Corresponding author}

@@

According to the latest mobility survey conducted in 2018 (INE, 2018),
the LMA registered a total of 5.3 million daily trips, with only 0.5\%
by bicycle. Car modal share was 58.4\%, while PT accounted for 15.5\%.
The number of intra-municipal trips --- with origin and destination in
the same municipality --- amounts to 3.5 million trips. This exceeds the
number of inter-municipal trips (1.8 million trips), involving travel
between different municipalities. Cars and public transport are the most
used modes for intercity trips, with cars being the predominant choice
for all journeys.

>>>>

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Modelling the impacts of replacing car trips with combined public transport and cycling: Reproducible methods, results and actionable evidence from the XXX project},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

  \title{Modelling the impacts of replacing car trips with combined
public transport and cycling: Reproducible methods, results and
actionable evidence from the XXX project}
    \author[CERIS]{Rosa Félix\corref{cor1}%
  \corref{cor1}%
  }
   \ead{rosamfelix@tecnico.ulisboa.pt} 
    \author[CERIS]{Filipe Moura%
  %
  }
   \ead{fmoura@tecnico.ulisboa.pt} 
    \author[ITS]{Robin Lovelace%
  %
  }
   \ead{r.lovelace@uleeds.uk} 
      \affiliation[CERIS]{CERIS, Instituto Superior Técnico -
Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal}
    \affiliation[ITS]{Institute for Transport Studies, University of
Leeds. 34-40 University Rd, Leeds LS2 9JT, UK}
    \cortext[cor1]{Corresponding author}

@@

According to the latest mobility survey conducted in 2018 (INE, 2018),
the LMA registered a total of 5.3 million daily trips, with only 0.5\%
by bicycle. Car modal share was 58.4\%, while PT accounted for 15.5\%
(see Figure \ref{fig:mododist}). The number of intra-municipal trips ---
with origin and destination in the same municipality --- amounts to 3.5
million trips. This exceeds the number of inter-municipal trips (1.8
million trips), involving travel between different municipalities. Cars
and public transport are the most used modes for intercity trips, with
cars being the predominant choice for all journeys.

\begin{figure}
\subfloat[In percentage\label{fig:mododist-1}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_files/figure-latex/mododist-1} }\subfloat[In total\label{fig:mododist-2}]{\includegraphics[width=0.5\linewidth,]{PaperCEUS_files/figure-latex/mododist-2} }\caption{\label{mododist}Trips in the LMA by inter/intra municipal and mode, according to the travel survey.}\label{fig:mododist}
\end{figure}
